---
title: "Data Visualization"
author: "Victor & Floriann"
format: 
  html: 
    code-fold: true
    embed-resources: true
    standalone: true
---

## Topic:
**04_TwoNumOrdered**
<br>
<p align='justify'>
1. Two numerical variables where one is ordered *(e.g.time)* and an additional categorical one *(Line plots, Area plots, Focusing, the Spaghetti plot problem)* (Inspiration from the stories 
</p>
<https://www.data-to-viz.com/story/TwoNumOrdered.html>, 
<br> <https://www.data-to-viz.com/story/OneCatSevOrderedNum.html>, and 
<br> <https://www.data-to-viz.com/caveat/spaghetti.html>)

<p align='justify'>
2. Tableau and Power BI
</p>

## Dataset
<p align='justify'>
1. Numerical variables where one is ordered (e.g. time) 
2. Categorical one 
</p>

```{r}
#| label: load-packages-and-data
#| message: false
library(tidyverse)
library(gapminder)
library(ggplot2)
library(babynames)
library(plotly)
```

**Dataset1: Bitcoin**

```{r}
#Data1
# Load dataset from github
bitcoin_data <- read.table("https://raw.githubusercontent.com/holtzy/data_to_viz/master/Example_dataset/3_TwoNumOrdered.csv", header=T)
bitcoin_data$date <- as.Date(bitcoin_data$date)
# Preview the data:
head(bitcoin_data)
```
<p align='justify'>
1. Numerical variable : `date`, `value`
</p>

**Dataset2: Babynames**

```{r}
#Data2
# Save the babynames data into baby_data:
babynames <- data.frame(babynames)
# Preview the data:
head(babynames)
```
<p align='justify'>
1. Numerical variable : `year`, `n`, `prop`
2. Categorical Variable : `sex`, `name` 
</p>

## Plots

1. **Line plot**

<p align='justify'>
Line chart depicts the evolution of one or several numeric variables. In other words, a line plot shows how the value of something changes over time.
</p>

```{r}
bitcoin_data %>%
  ggplot( aes(x=date, y=value)) +
    geom_line(color="#69b3a2") +
    ggtitle("Line plot - Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_light()
```

```{r}
# Load dataset from github
babynames <- babynames %>% 
  filter(name %in% c("Ashley", "Amanda", "Jessica",    "Patricia", "Linda", "Deborah",   "Dorothy", "Betty", "Helen")) %>%
  filter(sex=="F")

babynames %>%
  ggplot( aes(x=year, y=n, group=name, color=name)) +
    geom_line() +
    #scale_color_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Line plot - Popularity of American names in the previous 30 years") +
    theme_light()
```
<p align='justify'>
*Effective for*
<br>
**Comparing lots of data** all at once
<br>
1. Showing **changes and trends** over time
<br>
2. Including important **context and annotation**
<br>
3. Displaying **forecast data** and uncertainty
<br>
4. Highlighting **anomalies** within and across data series
</p>
<p align='justify'>
*Not the ideal choice for*
<br>
1. Working with categorical data
<br>
Showing sparse data sets
<br>
3. It does not tell us anything about the causes behind the fluctuation in the data.
<br>
4. Quite confusing when there is larger number of variables.
</p>

2. **Area plot**
<p align='justify'>
An area graph is a specialized form of the line graph, where instead of simply connecting our data points with a continuous line, we also fill in the region below that line with a solid color. 
</p>
<p align='justify'>

```{r}
bitcoin_data %>%
  ggplot( aes(x=date, y=value)) +
    geom_area(fill="#69b3a2", alpha=0.5) +
    geom_line(color="#69b3a2") +
    ggtitle("Area chart - Evolution of Bitcoin price") +
    ylab("bitcoin price ($)") +
    theme_light()
```

```{r}
babynames %>%
  ggplot( aes(x=year, y=n, group=name, fill=name)) +
    geom_area() +
    #scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Area chart - Popularity of American names in the previous 30 years") +
    theme_classic() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    facet_wrap(~name, scale="free_y")
```

```{r}
p <- babynames %>% 
  ggplot( aes(x=year, y=n, fill=name, text=name)) +
    geom_area( ) +
    #scale_fill_viridis(discrete = TRUE) +
    theme(legend.position="none") +
    ggtitle("Stack area chart - Popularity of American names in the previous 30 years") +
      theme() +
    theme(legend.position="none")
ggplotly(p, tooltip="text")
```

*Effective for*
<br>
1. Showing the **rise and fall** of various data series over time
<br>
2. Emphasizing a **part-to-whole relationship over time** when one part is very large, or changes from being very large to very small
<br>
3. Showing **change over time in individual panels** of a small multiple chart
</p>
<p align='justify'>
*Not the ideal choice for*
<br>
1. Data sets on scales that **do not have a meaningful relationship to zero**
<br>
2. Showing several **volatile data sets over time**
<br>
3. Showing **fine differences** in values
</p>

3. **Spaghetti plot**
<p align='justify'>
A Spaghetti plot is a line plot with many lines displayed together. It is a plot that depicts the outcome as a function of time for each individual subject.
</p>

```{r}
babynames %>%
  mutate( highlight=ifelse(name=="Ashley", "Ashley", "Other")) %>%
  ggplot( aes(x=year, y=n, group=name, color=highlight, size=highlight)) +
    geom_line() +
    scale_color_manual(values = c("#69b3a2", "lightgrey")) +
    scale_size_manual(values=c(1.5,0.2)) +
    theme(legend.position="none") +
    ggtitle("Spaghetti plot- Popularity of American names in the previous 30 years") +
    theme_light() +
    geom_label( x=1990, y=55000, label="Ashley reached 3550\nbabies in 1970", size=4, color="#69b3a2")
```
<p align='justify'>
*Effective for*
<br>
1. Maping the area you want to analyze
</p>
<p align='justify'>
*Problems with Spaghetti plot*
<br>
1. Hard to read, and thus provides few insights about the data.
</p>
<br>
[![Spaghetti plot in forecasting](https://i.insider.com/59b063a4609c3020008b46eb?width=1300&format=jpeg&auto=webp)](https://www.businessinsider.com/hurricane-irma-spaghetti-models-landfall-2017-9)
<br>
*Solutions:* <https://www.storytellingwithdata.com/blog/2013/03/avoiding-spaghetti-graph>

## Power BI and Tableau:

## References:
<p align='justify'>
1. https://allthingsstatistics.com/miscellaneous/line-graphs-advantages-disadvantages/
2. https://www.indeed.com/career-advice/career-development/a-guide-to-area-charts
3. https://www.storytellingwithdata.com/blog/2013/03/avoiding-spaghetti-graph
</p>